#Figure 1 - map with pie charts showing species recorded by iNat only, RLS only or Both Datasets



```{r}
library(tidyverse)
library(ggplot2)

#install.packages("scatterpie")
library(scatterpie)

library(ggrepel)
library(sf)


# Calculate species recorded by both iNat and RLS
iNat_RLS <- read.csv("Data/iNat_RLS.csv", header=TRUE)

iNat_RLS_SumSite_Bth  <- iNat_RLS  %>% 
    select(Species, Pres, Site, Dataset)  %>% 
    distinct()  %>% 
    group_by(Site, Species) %>%
    summarise(Pres = sum(Pres))

iNat_RLS_SumSite_Bth$Dataset <- "Both"
iNat_RLS_SumSite_Bth$Pres <- sub("1","0",iNat_RLS_SumSite_Bth$Pres)
iNat_RLS_SumSite_Bth$Pres <- sub("2","1",iNat_RLS_SumSite_Bth$Pres)

iNat_red_exYr  <- iNat_RLS  %>% 
  filter(Dataset=="iNat")  %>% 
    select(Species, Pres, Site, Dataset)  %>% 
    distinct() 

RLS_red_exYr  <- iNat_RLS  %>% 
  filter(Dataset=="RLS")  %>% 
    select(Species, Pres, Site, Dataset)  %>% 
    distinct() 

iNat_RLS_SumSite_Bth$Pres <- as.numeric(iNat_RLS_SumSite_Bth$Pres)

iNat_RLS_SumSite_Tot <- bind_rows(iNat_RLS_SumSite_Bth, iNat_red_exYr, RLS_red_exYr)

iNat_RLS_SumSite_Sum <- iNat_RLS_SumSite_Tot  %>% 
    group_by(Site, Dataset) %>%
    summarise(Total = sum(Pres))

iNat_RLS_SumSite_Sum_Spread <- spread(iNat_RLS_SumSite_Sum, Dataset, Total)
iNat_RLS_SumSite_Sum_Spread[is.na(iNat_RLS_SumSite_Sum_Spread)] <- 0
iNat_RLS_SumSite_Sum_Spread$iNat_unique <- iNat_RLS_SumSite_Sum_Spread$iNat-iNat_RLS_SumSite_Sum_Spread$Both
iNat_RLS_SumSite_Sum_Spread$RLS_unique <- iNat_RLS_SumSite_Sum_Spread$RLS-iNat_RLS_SumSite_Sum_Spread$Both
iNat_RLS_SumSite_Sum_unq <- select(iNat_RLS_SumSite_Sum_Spread, Site,  Both, RLS_unique, iNat_unique)
iNat_RLS_SumSite_Sum_unq <- gather(iNat_RLS_SumSite_Sum_unq, Both, RLS_unique, iNat_unique, key=Dataset, value = Total)
iNat_RLS_SumSite_Sum_unq$Dataset<-factor(iNat_RLS_SumSite_Sum_unq$Dataset, levels = c("iNat_unique", "RLS_unique", "Both"))

iNat_RLS_SumSite_Sum_unq$Year <- "Total"

iNat_RLS_SumSite_Sum_unq$Site <- factor(iNat_RLS_SumSite_Sum_unq$Site, levels = c("Clifton Gardens", "Kurnell", "Shiprock","Shelly Beach", "Camp Cove", "Bare Island",  "Gordons Bay", "Oak Park"))

# create the map


aus <- read_sf("Data/aust_cd66states.shp")

#random points dataframe to cover the random stars plotting on the map.. channel markers?
blockout <- data.frame(Lat=c(-33.839462, -33.9914, -33.832),
           Long=c(151.315, 151.185, 151.292))

site_coords <- data.frame(Site=c("Camp Cove", "Bare Island", "Shelly Beach", "Kurnell", "Shiprock", "Clifton Gardens", "Gordons Bay", "Oak Park"),
           Lat=c(-33.839462, -33.986114, -33.800491, -34.004803, -34.069222, -33.840133,-33.91537,-34.07022),
           Long=c(151.278486, 151.232260, 151.297299, 151.225942, 151.129858, 151.253429, 151.2674, 151.159))



iNat_RLS_SumSite_Sum_unq_coord <- left_join(iNat_RLS_SumSite_Sum_unq, site_coords)


iNat_RLS_SumSite_Sum_unq_coord_wide <- pivot_wider(iNat_RLS_SumSite_Sum_unq_coord, names_from=Dataset, values_from=Total)
iNat_RLS_SumSite_Sum_unq_coord_wide$Total <-  iNat_RLS_SumSite_Sum_unq_coord_wide$Both+iNat_RLS_SumSite_Sum_unq_coord_wide$RLS_unique+iNat_RLS_SumSite_Sum_unq_coord_wide$iNat_unique

rls_inat_syd_map_pie <- ggplot()+
  geom_sf(data=aus, fill="gray98")+
  theme_bw()+
  #random points to block stars/triangles off map
  geom_point(aes(x=Long, y=Lat), data=blockout, colour='white', size=4)+
    
  geom_scatterpie(aes(x=Long, y=Lat, group=Site, r=Total/12000), data=iNat_RLS_SumSite_Sum_unq_coord_wide,                                
                 cols=c("Both", "iNat_unique", "RLS_unique"), color=NA, alpha = 0.92) + 
theme(panel.grid.major=element_line(color="transparent"))+
  xlab("")+
  ylab("")+
  xlim(151.0, 151.4)+
  ylim(-34.1, -33.7)+
  coord_sf()+
 geom_text(data = iNat_RLS_SumSite_Sum_unq_coord_wide, aes(x = Long, y = Lat, label = Site),nudge_x = c(0.07, 0.06, 0.095, 0.065, 0.037, 0.05, 0.059, 0.067), 
                nudge_y = c(0 ,0, -0.02, 0, -0.01,0,0.03,0.019))+
   theme(legend.position = c(0.85, 0.13), legend.title = element_blank())+
   theme(text = element_text(size=12))+
  scale_fill_manual(values=c('#656565',"#A2E7F2", "#F98C3E"),labels=c ('Both', 'iNaturalist', 'RLS'))
 

rls_inat_syd_map_pie

#Australian map added outside R
```



#Figure 2 - species richness per year & species accumulation curve

```{r}
# Caclculate number of species per year
iNat_RLS_Comm.Fac_12yrs <- read.csv("Data/iNat_RLS_Comm_Fac_12yrs.csv")

iNat_RLS_sppR_12yrs <-iNat_RLS_Comm.Fac_12yrs


iNat_RLS_sppR_12yrs$SppR <- rowSums(iNat_RLS_sppR_12yrs[,4:457] )
iNat_RLS_sppR_12yrs <- select(iNat_RLS_sppR_12yrs, Year, Site, Dataset,  SppR)
iNat_RLS_sppR_12yrs$Site_Data <- paste(iNat_RLS_sppR_12yrs$Site, iNat_RLS_sppR_12yrs$Dataset, sep = "_")

# calculate species accumulation curves

library(BiodiversityR)

iNat_RLS_Comm_FacTab <- iNat_RLS_Comm.Fac_12yrs[1:3]
iNat_RLS_Comm_FacTab$SppR <- iNat_RLS_sppR_12yrs$SppR
iNat_RLS_Comm_FacTab$Site_Data <- paste(iNat_RLS_Comm_FacTab$Site, iNat_RLS_Comm_FacTab$Dataset, sep = "_")


#Add years with no data to factors table
iNat_RLS_Comm_allYrs_FacTab<- spread(iNat_RLS_Comm_FacTab,  Year,SppR)
iNat_RLS_Comm_allYrs_FacTab[is.na(iNat_RLS_Comm_allYrs_FacTab)] <- 0
iNat_RLS_Comm_allYrs_FacTab <- gather(iNat_RLS_Comm_allYrs_FacTab,  key=Year, value = SppR,'2008', '2009','2010','2011', '2012','2013', '2014', '2015','2016', '2017', '2018','2019')
iNat_RLS_Comm_allYrs_FacTab$Year <- as.numeric(iNat_RLS_Comm_allYrs_FacTab$Year)

#Add years with no data to community table 
iNat_RLS_Comm.Fac_allYrs <- right_join(iNat_RLS_Comm.Fac_12yrs, iNat_RLS_Comm_allYrs_FacTab)
iNat_RLS_Comm.Fac_allYrs[is.na(iNat_RLS_Comm.Fac_allYrs)] <- 0 

iNat_RLS_Comm_allYrs <- iNat_RLS_Comm.Fac_allYrs[4:457]
iNat_RLS_Comm_allYrs_FacTab <- iNat_RLS_Comm.Fac_allYrs[,c(1,2,3,458,459)]




Spp_accum_allYrs_coll <- accumcomp(iNat_RLS_Comm_allYrs, y=iNat_RLS_Comm_allYrs_FacTab, factor='Site_Data', xlab="Year",  method="collector", legend=FALSE)


Spp_accum_allYrs_coll_gg <- as.data.frame(Spp_accum_allYrs_coll)
Spp_accum_allYrs_coll_gg <- add_rownames(Spp_accum_allYrs_coll_gg, "Si_So")

Spp_accum_allYrs_coll_gg <- Spp_accum_allYrs_coll_gg %>%
  gather(key, value, -Si_So) %>%
  separate(key,into=c("Year","Data"))%>%
  spread(Data, value)%>%
  separate(Si_So,into=c("Site","Dataset"), sep="_")

Spp_accum_allYrs_coll_gg$Site <- factor(Spp_accum_allYrs_coll_gg$Site, levels = c("Clifton Gardens", "Shiprock", "Kurnell","Shelly Beach", "Camp Cove", "Bare Island", "Fairlight","Gordons Bay", "Oak Park"))

Spp_accum_allYrs_coll_gg$Year<-as.numeric(Spp_accum_allYrs_coll_gg$Year)  
 Spp_accum_allYrs_coll_gg$Year <- (Spp_accum_allYrs_coll_gg$Year + 2007)
Spp_accum_allYrs_coll_gg$Year<-as.character(Spp_accum_allYrs_coll_gg$Year)  

iNat_RLS_sppR_12yrs$Year<-as.character(iNat_RLS_sppR_12yrs$Year)  

# create combined plot of sppR/Yr and species accumulation curves

ggplot() +
  theme_bw()+
  theme(panel.grid = element_line(colour = "NA"))+
   geom_col(data=iNat_RLS_sppR_12yrs, aes(Year, SppR, fill= Dataset, group = Dataset), width=0.7,position=position_dodge(preserve='single'))+
    facet_wrap(~ Site, ncol = 2) +
  geom_line(data=Spp_accum_allYrs_coll_gg, aes(Year, Richness, colour = Dataset, group = Dataset), size=0.7) +
  scale_colour_manual(values=c("#A2E7F2", "#F98C3E"),name  ="Dataset",guide='legend', labels=c("iNaturalist", "RLS"))+
  labs(y= "Species Richness") + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.4)) +
  theme(text = element_text(size=12))+
 scale_y_continuous(breaks=seq(0, 400, 50), expand = c(0, 0), limits = c(0, 400)) +
   theme(legend.position = c(0.856, 0.94), legend.title = element_blank())+
  #scale_x_continuous(breaks=seq(2017, 2019, 1))+
    scale_fill_manual(values=c("#A2E7F2", "#F98C3E"),name  ="Dataset",guide='legend', labels=c("iNaturalist", "RLS"))

```




#Section 3.2 species composition 
# Supplimentary Material 3 - GLLVM ordination plot
```{r}


#Create  community matrix including Factors
iNat_RLS_Comm.Fac <- spread(iNat_RLS, Species, Pres, fill = 0, convert = FALSE, drop = TRUE,
  sep = NULL)
#Create Site_Year_Dataset Row label
iNat_RLS_Comm <- unite(iNat_RLS_Comm.Fac, Site, Year, Dataset, col="Site_Yr_Data",sep="_")
iNat_RLS_Comm<- column_to_rownames(iNat_RLS_Comm, var = "Site_Yr_Data")



# gllvm test


library(gllvm)

iNat_RLS_gllvm <-gllvm(iNat_RLS_Comm,num.lv=2,family=binomial(link=cloglog), row.eff="random")


# Check Assumptions - residuals
plot (iNat_RLS_gllvm)


# Create ordination plot - Supplementary Material 3


gllvm_lvls<-as.data.frame(iNat_RLS_gllvm$lvs) 

gllvm_lvls<- cbind(gllvm_lvls,iNat_RLS_Comm.Fac$Site)
gllvm_lvls<- cbind(gllvm_lvls,iNat_RLS_Comm.Fac$Dataset)
gllvm_lvls <- rename(gllvm_lvls, Site='iNat_RLS_Comm.Fac$Site')
gllvm_lvls <- rename(gllvm_lvls,Dataset='iNat_RLS_Comm.Fac$Dataset')

ggplot(gllvm_lvls, aes(x=LV1, y=LV2, color = iNat_RLS_Comm.Fac$Dataset,shape=iNat_RLS_Comm.Fac$Site)) + 
  geom_point(size=4.5, stroke = 1.5) + 
  theme_bw() +
  theme(panel.grid = element_blank())+
  theme(legend.key.size = unit(.9, "cm")) +
    scale_shape_manual(values = c(3,18,7, 16, 4,8,17,15), name  ="Site")  + 
  scale_colour_manual(values=c("#A2E7F2", "#F98C3E"),name  ="Dataset",guide='legend', labels=c("iNaturalist", "RLS"))
```

#testing differences in community composition - mvabund test - Section 3.2 
```{r}
library(mvabund)


iNat_RLS_mvabund <-  mvabund(iNat_RLS_Comm.Fac[,4:374])


iNat_RLS_Comm_Facs <- iNat_RLS_Comm.Fac[,1:3]
iNat_RLS_Comm_Facs$Dataset <- as.factor(iNat_RLS_Comm_Facs$Dataset)
iNat_RLS_Comm_Facs$Site <- as.factor(iNat_RLS_Comm_Facs$Site)
iNat_RLS_Comm_Facs$Year <- as.numeric(iNat_RLS_Comm_Facs$Year)


#Many GLM by Site

iNat_RLS_manyglm_SixDa_Yr <- manyglm(iNat_RLS_mvabund~Site*Dataset+Year, data=iNat_RLS_Comm_Facs, family=binomial("cloglog"))


iNat_RLS_SixDa_Yr_ANOVA <- anova(iNat_RLS_manyglm_SixDa_Yr, p.uni="adjusted", test="LR", cor.type = "I", nBoot=1000)


iNat_RLS_SixDa_Yr_ANOVA

```

#Comparison of datasets for each site - pairwise tests

```{r}
Shelly_Comm.Fac <- filter(iNat_RLS_Comm.Fac, Site =='Shelly Beach')
Shelly_Facs <- Shelly_Comm.Fac[1:3]
Shelly_Comm.Fac.subset = Shelly_Comm.Fac[4:374]
Shelly_Comm.Fac.colSum = apply(Shelly_Comm.Fac.subset , 2, sum)
Shelly_Comm.Fac.subset = Shelly_Comm.Fac.subset [,which(Shelly_Comm.Fac.colSum >0)]
Shelly_mvabund <-  mvabund(Shelly_Comm.Fac.subset)

Shelly_manyglm <- manyglm(Shelly_mvabund~Dataset + Year, data=Shelly_Facs, family=binomial("cloglog"))

Shelly_ANOVA <- anova(Shelly_manyglm,  nBoot=1000, test="LR", cor.type = "I")

Shelly_ANOVA_uni <- anova(Shelly_manyglm,  nBoot=1000,  p.uni="adjusted", test="LR", cor.type = "I")


Bare_Comm.Fac <- filter(iNat_RLS_Comm.Fac, Site =='Bare Island')
Bare_Facs <- Bare_Comm.Fac[1:3]
Bare_Comm.Fac.subset = Bare_Comm.Fac[4:374]
Bare_Comm.Fac.colSum = apply(Bare_Comm.Fac.subset , 2, sum)
Bare_Comm.Fac.subset = Bare_Comm.Fac.subset [,which(Bare_Comm.Fac.colSum >0)]
Bare_mvabund <-  mvabund(Bare_Comm.Fac.subset)

Bare_manyglm <- manyglm(Bare_mvabund~Dataset+ Year, data=Bare_Facs, family=binomial("cloglog"))

Bare_ANOVA <- anova(Bare_manyglm,  nBoot=1000, test="LR", cor.type = "I")



Camp_Comm.Fac <- filter(iNat_RLS_Comm.Fac, Site =='Camp Cove')
Camp_Facs <- Camp_Comm.Fac[1:3]
Camp_Comm.Fac.subset = Camp_Comm.Fac[4:374]
Camp_Comm.Fac.colSum = apply(Camp_Comm.Fac.subset , 2, sum)
Camp_Comm.Fac.subset = Camp_Comm.Fac.subset [,which(Camp_Comm.Fac.colSum >0)]
Camp_mvabund <-  mvabund(Camp_Comm.Fac.subset)

Camp_manyglm <- manyglm(Camp_mvabund~Dataset+ Year, data=Camp_Facs, family=binomial("cloglog"))

Camp_ANOVA <- anova(Camp_manyglm,  nBoot=1000, test="LR", cor.type = "I")




Clifton_Comm.Fac <- filter(iNat_RLS_Comm.Fac, Site =='Clifton Gardens')
Clifton_Facs <- Clifton_Comm.Fac[1:3]
Clifton_Comm.Fac.subset = Clifton_Comm.Fac[4:374]
Clifton_Comm.Fac.colSum = apply(Clifton_Comm.Fac.subset , 2, sum)
Clifton_Comm.Fac.subset = Clifton_Comm.Fac.subset [,which(Clifton_Comm.Fac.colSum >0)]
Clifton_mvabund <-  mvabund(Clifton_Comm.Fac.subset)

Clifton_manyglm <- manyglm(Clifton_mvabund~Dataset+ Year, data=Clifton_Facs, family=binomial("cloglog"))

Clifton_ANOVA <- anova(Clifton_manyglm,  nBoot=1000, test="LR", cor.type = "I")



Gordons_Comm.Fac <- filter(iNat_RLS_Comm.Fac, Site =='Gordons Bay')
Gordons_Facs <- Gordons_Comm.Fac[1:3]
Gordons_Comm.Fac.subset = Gordons_Comm.Fac[4:374]
Gordons_Comm.Fac.colSum = apply(Gordons_Comm.Fac.subset , 2, sum)
Gordons_Comm.Fac.subset = Gordons_Comm.Fac.subset [,which(Gordons_Comm.Fac.colSum >0)]
Gordons_mvabund <-  mvabund(Gordons_Comm.Fac.subset)

Gordons_manyglm <- manyglm(Gordons_mvabund~Dataset+ Year, data=Gordons_Facs, family=binomial("cloglog"))

Gordons_ANOVA <- anova(Gordons_manyglm,  nBoot=1000, test="LR", cor.type = "I")




Oak_Comm.Fac <- filter(iNat_RLS_Comm.Fac, Site =='Oak Park')
Oak_Facs <- Oak_Comm.Fac[1:3]
Oak_Comm.Fac.subset = Oak_Comm.Fac[4:374]
Oak_Comm.Fac.colSum = apply(Oak_Comm.Fac.subset , 2, sum)
Oak_Comm.Fac.subset = Oak_Comm.Fac.subset [,which(Oak_Comm.Fac.colSum >0)]
Oak_mvabund <-  mvabund(Oak_Comm.Fac.subset)

Oak_manyglm <- manyglm(Oak_mvabund~Dataset+ Year, data=Oak_Facs, family=binomial("cloglog"))

Oak_ANOVA <- anova(Oak_manyglm,  nBoot=1000, test="LR", cor.type = "I")



Shiprock_Comm.Fac <- filter(iNat_RLS_Comm.Fac, Site =='Shiprock')
Shiprock_Facs <- Shiprock_Comm.Fac[1:3]
Shiprock_Comm.Fac.subset = Shiprock_Comm.Fac[4:374]
Shiprock_Comm.Fac.colSum = apply(Shiprock_Comm.Fac.subset , 2, sum)
Shiprock_Comm.Fac.subset = Shiprock_Comm.Fac.subset [,which(Shiprock_Comm.Fac.colSum >0)]
Shiprock_mvabund <-  mvabund(Shiprock_Comm.Fac.subset)

Shiprock_manyglm <- manyglm(Shiprock_mvabund~Dataset+ Year, data=Shiprock_Facs, family=binomial("cloglog"))

Shiprock_ANOVA <- anova(Shiprock_manyglm,  nBoot=1000, test="LR", cor.type = "I")


Kurnell_Comm.Fac <- filter(iNat_RLS_Comm.Fac, Site =='Kurnell')
Kurnell_Facs <- Kurnell_Comm.Fac[1:3]
Kurnell_Comm.Fac.subset = Kurnell_Comm.Fac[4:374]
Kurnell_Comm.Fac.colSum = apply(Kurnell_Comm.Fac.subset , 2, sum)
Kurnell_Comm.Fac.subset = Kurnell_Comm.Fac.subset [,which(Kurnell_Comm.Fac.colSum >0)]
Kurnell_mvabund <-  mvabund(Kurnell_Comm.Fac.subset)

Kurnell_manyglm <- manyglm(Kurnell_mvabund~Dataset+ Year, data=Kurnell_Facs, family=binomial("cloglog"))

Kurnell_ANOVA <- anova(Kurnell_manyglm,  nBoot=1000, test="LR", cor.type = "I")



Shelly_ANOVA
Bare_ANOVA
Camp_ANOVA
Clifton_ANOVA
Gordons_ANOVA
Oak_ANOVA
Shiprock_ANOVA
Kurnell_ANOVA
```




# Figure 3 - Bi-directional Bar plots of species occurence
Based on significant species from ANOVA univariate tests 

```{r}
iNat_RLS_SixDa_ANOVA_uni <- as.data.frame(iNat_RLS_SixDa_Yr_ANOVA$uni.p)
iNat_RLS_SixDa_ANOVA_uni <- iNat_RLS_SixDa_ANOVA_uni[-1, ]

iNat_RLS_SixDa_ANOVA_uni <- rownames_to_column(iNat_RLS_SixDa_ANOVA_uni, var="Variable")


colnames(iNat_RLS_SixDa_ANOVA_uni) <- gsub("\\.", " ", colnames(iNat_RLS_SixDa_ANOVA_uni))


iNat_RLS_SppTot_Dataset <- iNat_RLS_Comm.Fac %>%
  select(3:370)  %>%
  group_by(Dataset) %>%
  summarise_all(sum)%>%
rename(Variable=Dataset)

colnames(iNat_RLS_SppTot_Dataset) <- gsub("\\.", " ", colnames(iNat_RLS_SppTot_Dataset))

#Join uni pvalues to count data
iNat_RLS_SixDa_ANOVA_uni_counts <-bind_rows(iNat_RLS_SixDa_ANOVA_uni,iNat_RLS_SppTot_Dataset)



iNat_RLS_SixDa_ANOVA_uni_counts <-column_to_rownames(iNat_RLS_SixDa_ANOVA_uni_counts, var="Variable")

iNat_RLS_SixDa_ANOVA_uni_counts <- as.data.frame(t(iNat_RLS_SixDa_ANOVA_uni_counts))


# Filter to Dataset <0.05 (i.e. significant diff between Datasets)
iNat_RLS_SixDa_ANOVA_uni_counts <-rownames_to_column(iNat_RLS_SixDa_ANOVA_uni_counts, var="Species")
iNat_RLS_SixDa_ANOVA_uni_sig_data  <- filter(iNat_RLS_SixDa_ANOVA_uni_counts, Dataset < 0.05)


# Conver iNat v RLS into +- Counts  for bi directionality of plot

iNat_RLS_SixDa_ANOVA_uni_sig_data <- select(iNat_RLS_SixDa_ANOVA_uni_sig_data, Species, iNat, RLS)
iNat_RLS_SixDa_ANOVA_uni_sig_data <- gather(iNat_RLS_SixDa_ANOVA_uni_sig_data, iNat, RLS, key=Dataset,value='Counts')


#bidirectional plot with fish silhouettes 
#unhash to install fisualize if required
#library(devtools)
#devtools::install_github("nschiett/fishualize", force = TRUE)
library(fishualize)

sig_species_bar_fishpics <- ggplot(data = iNat_RLS_SixDa_ANOVA_uni_sig_data, 
       mapping = aes(x = reorder(Species, Counts, diff), fill = Dataset, 
                     y = ifelse(test = Dataset == "RLS", 
                                yes = -Counts, no = Counts))) +
  geom_bar(stat = "identity") +
  scale_y_continuous(labels = abs, limits = max(iNat_RLS_SixDa_ANOVA_uni_sig_data$Counts) * c(-1,1.2)) +
  labs(y = "Number of Records") +
  coord_flip() +
  theme_bw() +
  theme(panel.grid = element_line(colour = "NA"), axis.text.y = element_text(face="italic", size=8),axis.text.x = element_text(size=12),axis.title.x = element_text(size=12), axis.title.y = element_blank()) +
  scale_fill_manual(values=c("#A2E7F2", "#F98C3E"), name  ="Dataset",breaks=c("iNat", "RLS"), labels=c("iNaturalist", "RLS"))+
  theme(text = element_text(size=12))+
   theme(legend.position = c(0.19, 0.15), legend.title = element_blank())+

### Code below adds fish shapes - may plot in wrong position depending on the number of significant species returned in ANOVA uni tests
  ### Some fish families were not loading correctly so were added manually. 
 add_fishape(family = "Tripterygiidae",
            option = "Helcogramma_fuscopinna",
              xmin = 17.5, xmax = 18.5, ymin = 5, ymax = 10,
              fill = "black",
              alpha = 1) +

add_fishape(family = "Pomacentridae",
             option = "Abudefduf_sexfasciatus",
             xmin = 16.5, xmax = 17.5, ymin = 8, ymax = 13,
              fill = "black",
              alpha = 1) +
  
#   add_fishape(family = "Cheilodactylidae",
#   option = "Cheilodactylus_fuscus",
#   xmin = 15.5, xmax = 16.5, ymin = 18, ymax = 22,
#   fill = "black",
#   alpha = 1)+
  
add_fishape(family = "Syngnathidae",
            option = "Syngnathus_sp",
                     xmin = 15.5, xmax = 16.5, ymin = 8, ymax = 16,
              fill = "black",
              alpha = 1) +

add_fishape(family = "Blenniidae",
            option = "Antennablennius_adenensis",
                     xmin = 14.5, xmax = 15.5, ymin = 9, ymax = 15,
              fill = "black",
              alpha = 1) +

##add_fishape(family = "Platycephalidae",
##   option = "Thysanophrys_cirronasa",
##                    xmin = 13.5, xmax = 14.5, ymin = 10, ymax = 16,
##              fill = "black",
##             alpha = 1) +
  
  add_fishape(family = "Tetraodontidae",
            option = "Arothron_meleagris",
                     xmin = 12.5, xmax = 13.5, ymin = 10, ymax = 16,
              fill = "black",
              alpha = 1) +

add_fishape(family = "Syngnathidae",
            option = "Syngnathus_sp",
                     xmin = 11.5, xmax = 12.5, ymin = 11, ymax = 19,
              fill = "black",
              alpha = 1) +
  
  add_fishape(family = "Syngnathidae",
            option = "Syngnathus_sp",
                     xmin = 10.5, xmax = 11.5, ymin = 11, ymax = 19,
             fill = "black",
              alpha = 1) +

    add_fishape(family = "Dasyatidae",
            option = "Bathytoshia_brevicaudata",
                     xmin = 9.5, xmax = 10.8, ymin = 11, ymax = 14,
              fill = "black",
              alpha = 1) +
  

add_fishape(family =  "Kyphosidae",
              option = "Kyphosus_cinerascens",
                     xmin = 8.2, xmax = 9.7, ymin = 20, ymax = 24,
              fill = "black",
              alpha = 1) +

add_fishape(family =  "Hypnidae",
   option = "Hypnos_monopterygius", 
                    xmin = 7.2, xmax = 8.8, ymin = 12, ymax = 15,
              fill = "black",
             alpha = 1) +

  add_fishape(family =  "Urolophidae",
 option = "Trygonoptera_testacea",
                     xmin = 6.2, xmax = 7.8, ymin = 13, ymax = 16,
              fill = "black",
             alpha = 1) +
  
    add_fishape(family =  "Labridae",
              option = "Coris_gaimard",
                     xmin = 5.5, xmax = 6.5, ymin = 13, ymax = 17,
              fill = "black",
              alpha = 1) +
  
# #   add_fishape(family =  "Scorpaenidae",
##            option = "Pterois_volitans",
# #                    xmin = 4.5, xmax = 6.5, ymin = 15, ymax = 25,
#  #            fill = "black",
#   #           alpha = 1) +
  
##  add_fishape(family = "Platycephalidae",
# #  option = "Thysanophrys_cirronasa",
#  #                  xmin = 3.5, xmax = 4.5, ymin = 16, ymax = 22,
#   #           fill = "black",
#    #         alpha = 1) +

# #   add_fishape(family = "Heterodontidae",
# #  option = "Heterodontus_portusjacksoni",
# #                   xmin = 2.5, xmax = 3.5, ymin = 16, ymax = 22,
# #            fill = "black",
# #           alpha = 1) +
      
      add_fishape(family = "Diodontidae",
   option = "Dicotylichthys_punctulatus",
                    xmin = 1.5, xmax = 2.5, ymin = 19, ymax = 24,
              fill = "black",
             alpha = 1) +
  
    #technically not the right family -> should be Aracanidae
    add_fishape(family =  "Ostraciidae",
              option = "Ostracion_cubicus",
                     xmin = 0.5, xmax = 1.5, ymin = 17, ymax = 23,
              fill = "black",
              alpha = 1) 
   
   
sig_species_bar_fishpics

```




#Figure 4 - iNaturalist photographic observations
```{r}
iNat <- read.csv("Data/iNat_Obs.csv", header=TRUE)

#iNat Photo count

iNat_Photo_Count <- select(iNat, Species, Pres, Site, Date)
iNat_Photo_Count$Date <- as.Date(iNat_Photo_Count$Date)
iNat_Photo_Count$Date <- format(iNat_Photo_Count$Date, "%Y")
iNat_Photo_Count <- filter(iNat_Photo_Count, Date>2007)
iNat_Photo_Count <- filter(iNat_Photo_Count, Date<2020)
iNat_Photo_Count$Site_Yr <- as.factor(paste(iNat_Photo_Count$Site, iNat_Photo_Count$Date, sep = "_"))


iNat_Photo_Count <- iNat_Photo_Count  %>% 
  group_by(Site_Yr) %>%
  summarise(Photo_Count = sum(Pres))

iNat_Photo_Count <- separate(iNat_Photo_Count, 1, c("Site", "Year"), sep="_")


ggplot(data=iNat_Photo_Count, aes(Year, Photo_Count, colour = Site, group = Site )) +
  theme_classic() +
  theme(panel.grid = element_line(colour = "NA"), text=element_text(size=12))+
     geom_line( size=0.5,show.legend = FALSE) +
  labs(y= "Photos per Year")+
  scale_y_log10()+
    scale_x_discrete(breaks = c(2008,2009,2010,2011,2012,2013,2014,2015,2016,2017,2018,2019), expand = c(0.03, 0.03))+
    theme(axis.text.x = element_text(angle = 90, vjust = 0.4)) +
theme(legend.position='bottom', legend.title = element_blank(),legend.justification = c(0.9,0.9))+
  geom_point(aes(shape=Site), size=3, stroke = 1)+
   theme(text = element_text(size=12))+
    scale_shape_manual(values = c(3,18,7, 16, 4,8,17,15), name  ='Site') 
```

# Figure 5 - Number of species per observation event
```{r}
RLS_M1 <- read.csv("Data/RLS_M1.csv", header=TRUE)
RLS_M2 <- read.csv("Data/RLS_M2.csv", header=TRUE)

RLS <- bind_rows(RLS_M1,RLS_M2)

#Create reduce RLS table
RLS_trans <- select(RLS, Species, Pres, Date,Depth, Site)

RLS_trans$Date <- as.Date(RLS_trans$Date)
RLS_trans$Pres <- as.integer(RLS_trans$Pres)


RLS_trans <- select(RLS_trans, Species, Pres, Date,Depth, Site)

#Remove duplicate species records for a given site and year
RLS_trans <- distinct(RLS_trans, .keep_all = FALSE)

#Create RLS community table including Factors
RLS_trans_inc.Fac <- spread(RLS_trans, Species, Pres, fill = 0, convert = FALSE, drop = TRUE,
  sep = NULL)

# Species/Transect
sp_trans<-RLS_trans_inc.Fac

sp_trans$Total <-  sp_trans %>%
  select (4:153) %>%
  rowSums()

#sp_trans$Year<- format(sp_trans$Date, "%Y")  
sp_trans$Year <- substr(sp_trans$Date, 0, 4)


sp_trans_tab <- sp_trans %>% 
  group_by(Site) %>% 
summarise(observations = sum (Total), species_transect = mean(Total), SD = sd(Total), transects=n(), min=min(Total), max=max(Total))

sp_trans_tab$SE <- sp_trans_tab$SD/(sqrt(sp_trans_tab$transects))

# Photographs/Dive

iNat_3yrs <- subset(iNat, Date>="2017-01-01" & Date <="2019-12-31")

Photos_dive <-  iNat_3yrs %>% 
  group_by(Site, Date, User) %>% 
summarise(Total = n())

#Photos_dive$Year <- format(Photos_dive$Date, "%Y")    
Photos_dive$Year <- substr(Photos_dive$Date, 0, 4)

trans_photos <- sp_trans[c(2:3,236:237)]
trans_photos$Dataset <- 'RLS'

Photos_dive$Dataset <- 'iNat'

trans_photos <- bind_rows(trans_photos,Photos_dive)


ggplot(trans_photos, aes(x = Total, fill=Dataset))+  
    theme_bw()+
  theme(panel.grid = element_line(colour = "NA"))+
  geom_histogram( aes(y = 1*..density.. ), binwidth = 1,  position = "identity") +
  facet_wrap(~ Dataset, ncol = 2) +
 scale_alpha_manual(values = c(1,  0.5)) +
  scale_colour_manual(values=c("#82bbc4","#cf7434")) +
    scale_x_continuous(breaks = c(0,10,20,30,40,50,60,70), expand = c(0.02,0.1)) +
  scale_y_continuous(breaks = c(0,0.01,0.05,.1,.2,.3,.4,.5), labels = paste0(c(0,0.01,0.05,.1,.2,.3,.4,.5)*100," %"), expand= c(0,0.008),  limits = c(0, 0.5), trans='sqrt') +
   theme(legend.position = "none")+  
   labs(y= "Proportion of Events", x="Species Recorded per Observation Event") +
     theme(text = element_text(size=12))+
  scale_fill_manual(values=c("#A2E7F2", "#F98C3E"),name  ="Dataset",guide='legend', labels=c("iNaturalist", "RLS") )


```

#Supplementary Material 2 - Average species richness plot 
```{r}
iNat_RLS_sppR_3yrs <-iNat_RLS_Comm.Fac
iNat_RLS_sppR_3yrs$SppDiv <- rowSums(iNat_RLS_sppR_3yrs[,4:374] )

avg_rich <- iNat_RLS_sppR_3yrs %>%
  group_by(Site, Dataset) %>%
  summarise(mean= mean(SppDiv), sd=sd(SppDiv), n = n_distinct(Year))

avg_rich$SE <- avg_rich$sd/sqrt(avg_rich$n)
avg_rich$Year <- "Average"
avg_rich <- rename(avg_rich, SppDiv=mean)
avg_rich <-select(avg_rich,-sd,-n)


my_breaks <- function(y) { if (max(y) < 100) seq(0, 100, 25) else  seq(0, 200, 50)}

ggplot() +
  theme_bw()+
  theme(panel.grid = element_line(colour = "NA"))+
   geom_boxplot(data=iNat_RLS_sppR_3yrs, aes(Dataset, SppDiv, fill= Dataset), width=0.3,position=position_dodge(preserve='single'))+
   facet_wrap(~ Site, ncol = 2, scales = "free_y")+
labs(y= "Species Richness") + 
  #theme(axis.text.x = element_text(angle = 90)) +
 expand_limits( y = 0)+
    #scale_y_continuous(breaks=seq(0, 200,50), expand = c(0, 0), limits = c(0, 200)) +
    scale_fill_manual(values=c("#A2E7F2", "#F98C3E"),name  ="Dataset", labels=c("iNaturalist", "RLS") ) +
scale_y_continuous(breaks = my_breaks)


```
#ANOVA/Pairwise test for difference in average annual species richness
```{r}
#install.packages("afex")
library(afex)



iNat_RLS_sppR_3yrs <- select(iNat_RLS_sppR_3yrs, Site, Dataset,  SppDiv)

iNat_RLS_sppR_3yrs <- rownames_to_column(iNat_RLS_sppR_3yrs, "ID")

iNat_RLS_sppR_3yrs$ID <- as.integer (iNat_RLS_sppR_3yrs$ID)

SppR_Anova<-aov_car(SppDiv~Site*Dataset  + Error(ID), 
                  data=iNat_RLS_sppR_3yrs)
SppR_Anova

#install.packages("emmeans")
library(emmeans)
Simple.Effects.By.Site<-emmeans(SppR_Anova, ~Dataset|Site)
Simple.Effects.By.Site



#as F-tests
test(pairs(Simple.Effects.By.Site), joint = TRUE)




```


#Supplementary Material 4 - Effort - Sampling events. Dives/transects

```{r}
#Create reduce RLS table
RLS_transect_year <- select(RLS, Date, Depth, Site)

#Remove duplicate species records for a given site and year
RLS_transect_year <- distinct(RLS_transect_year, .keep_all = FALSE)

    RLS_transect_year$Date <- as.Date(RLS_transect_year$Date)
RLS_transect_year$Year <- format(RLS_transect_year$Date, "%Y")
RLS_transect_year <- RLS_transect_year %>%
  group_by(Site, Year) %>%
   count(name="Transects")

RLS_transect_year <- rename(RLS_transect_year, Total=Transects)
 RLS_transect_year$Dataset <- 'RLS'
 RLS_transect_year$Year <- as.numeric(RLS_transect_year$Year)   

iNat_dives_yr <- iNat
iNat_dives_yr$Year <- substr(iNat_dives_yr$Date, 0, 4)
iNat_dives_yr$Year <- as.numeric(iNat_dives_yr$Year)   
iNat_dives_yr <-  subset(iNat_dives_yr, Year>=2008 & Year <=2019)


  iNat_dives_yr <-  iNat_dives_yr %>% 
  distinct(Year, Site, User, .keep_all=TRUE)%>% 
  group_by(Site, Year) %>% 
summarise(Total = n())
  
  iNat_dives_yr$Dataset <- 'iNat'
  
  

 effort_events <- rbind(iNat_dives_yr,RLS_transect_year )
 
ggplot() +
  theme_bw()+
  theme(panel.grid = element_line(colour = "NA"))+
    facet_wrap(~ Site, ncol = 2) +
  geom_col(data=effort_events, aes(Year, Total, fill = Dataset, group = Dataset), size=1,position=position_dodge(preserve='single')) +
  scale_fill_manual(values=c("#A2E7F2", "#F98C3E"),name  ="Dataset",guide='legend', labels=c("iNaturalist", "RLS"))+
  labs(y= "Number of Events") + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.4)) +
 scale_y_continuous(breaks=seq(0, 30, 5), expand = c(0, 0), limits = c(0, 30)) +
  scale_x_continuous(breaks=seq(2008, 2019, 1))

```







